\ifdefined\MAINTEXT \else Supplemental Material: \\ \fi Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2017-10-02 almo-langevin.tex
No abstract

\author{Hayden Scheiber}
\email{hayden.scheiber@mcgill.ca}
\author{Yifei Shi}
\author{Rustam Z. Khaliullin}
\email{rustam.khaliullin@mcgill.ca}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}

@@

\begin{abstract}
A density functional theory approach based on absolutely localized molecular orbitals is used in conjunction with the Langevin equation of motion to create a low-cost molecular dynamics method for weakly coupled molecular systems which retains sufficient accuracy to generate accurate dynamical properties in water.

%The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
%The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
%In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.

@@

%RZK-important addition: we did not include a large family of the so-called fragmentation methods 
%included only those methods that attempt to construct  set 
%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix. This approach provides the most rigorous description of the electronic structure.

@@

Once the localization radius is chosen, ALMO constraint prohibits electron density transfer between distant molecules but retains all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.

%To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to compensate errors in the forces~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}. 
%We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
%We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.

The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals. First, while ALMOs are variationally optimized in both SCF steps, the occupied-space projector obtained in the first step must remain fixed during the second step to ensure convergence. Second, some neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$. In addition to the ALMO-specific issues, the variational optimization in any AIMD methods is never truly complete and is interrupted once the norm of the gradient of the energy with respect to the variational electronic degrees of freedom drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$. While these non-SCF errors are minor and do not affect accuracy of static ALMO DFT calculations, geometry optimization, and ALMO-based Monte-Carlo simulations, they accumulate in molecular dynamics trajectories resulting in non-physical sampling and eventually disrupting them.

%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated. 
%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{RZK-any-references}. This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.

%While these issues can be resolved by  by increasing  to account for non-SCF forces, increasing  and decreasing While  are the known formal strategies for resolving these

In order to maintain stable and accurate dynamics, one could use a set of strategies that includes computing the non-SCF contribution to the forces through a coupled perturbed procedure, increasing $R_{c}$, and decreasing $\epsilon_{\text{SCF}}$. %relative to the corresponding optimal values for static energy calculations or Monte-Carlo simulations. 
This however would dramatically increase the computational cost of the ALMO AIMD. %as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver. 
%
In this work we propose another approach, which obviates the need in a coupled-perturbed solver, allows for relaxation of the tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus keeps the complexity and cost of simulations low.

In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense non-SCF component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that would be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$. %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
%RZK For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}. 
Inspired by this method, ALMO AIMD is chosen by be governed by the Langevin equations of motion, which can be written using the unknown reference forces as:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is stochastic force represented by a white Gaussian noise with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
%
This relation means that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms will maintain a constant average temperature $T$ and generate the correct Maxwell-Boltzmann distribution of velocities~\cite{a:Kubo-1986,b:tuckerman-stat}. 
In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled. 
Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to more random noise in trajectories, faster equilibration but more distorted results for calculated dynamical properties.

@@

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown below to be well justified, allows us to rewrite the Langevin equations using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for the imperfection of ALMO forces and is required to re-balance the damping and stochastic components in the Langevin equation and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.

@@

%To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
%This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
%This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
%Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.

(RZK the integral contains dummy index $i$ in the r.h.s.) In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle \delta f_{i\alpha} (0)  \delta f_{i\alpha} (t) \rangle dt
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory (see Supplemental Material). 
In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. 
This or, in fact, any starting value can be  rapidly fine-tuned in a series of trial-and-error short ALMO AIMD runs until the equipartition theorem is satisfied: 
%
\begin{align}
\label{eq:eqipartition}
\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
\end{align}
%
or, in other words, until the simulation stabilizes with the correct average temperature.

@@

ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly in approximate ALMO AIMD to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{298}{\K}$ and a constant density of $\SI{1.006}{\g\cdot\cm^{-3}}$. 
% RZK: we do not need to specify details on gamma optimization since they are well-known.
RZK algorithm~\cite{RZK} was used to integrate the Langevin equation of motion. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system on the time scale of the simulation and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual-representation Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
The Brillouin zone was sampled at the $\Gamma$-point and separable norm-conserving pseudopotentials were used to describe the interactions between the valence electrons and the ionic cores~\cite{a:hgh}.

The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot,a:ot2}. 
In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR). This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the KS reference RDF perfectly in Monte-Carlo simulations~\cite{a:almo-ls}. 
%
To measure the ability of the $R^{\Delta}$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.

%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature. The fact the the actual temperature deviates from the target temperature by 5 K even with perfect forces shows that our simulation is still too short.
%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
%All other system parameters were set identical to the ALMO system.

Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b). $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$. 
%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because unnecessary stochastic noise is eliminated in the trial runs. %In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
\caption{\label{fig:dynproperties} RZK:. Calculated properties of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) radial distribution function, 
(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) infrared spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

Analysis of $\delta \vec{f_{i}}(t)$ shows that the error in forces indeed resembles Gaussian white noise. The mean of the error is zero (black line in Figure~\ref{fig:randomforce}c) and the ACF decays rapidly (Figure~\ref{fig:randomforce}d) so that the error in forces can be considered uncorrelated on time scale of $\SI{50}{\fs}$. Thus the main assumptions behind our approach to ALMO AIMD are justified for liquid water. Figure~\ref{fig:randomforce}b shows that the main source of error in forces in this simulation is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error. We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ (Figure~\ref{fig:forcecomp} in the Supplemental Material).

@@

%RZK Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.

>>>>

\author{Hayden Scheiber}
% RZK update email
\email{hayden.scheiber@mcgill.ca}
\author{Yifei Shi}
\author{Rustam Z. Khaliullin}
\email{rustam.khaliullin@mcgill.ca}
\affiliation{Department of Chemistry, McGill University, 801 Sherbrooke St. West, Montreal, QC H3A 0B8, Canada}

@@

\begin{abstract} 
%The utility of density functional theory based on compact localized molecular orbitals in \emph{ab initio} molecular dynamics. To overcome limitations of existing linear-scaling methods
%%in conjunction with a modified Langevin equation of motion 
%to create a linear-scaling \emph{ab initio} molecular dynamics method for weakly coupled molecular systems. 
%To overcome %which retains sufficient accuracy to generate accurate dynamical properties in water. 
%%Strictly localized 
%%Low-cost linear-scaling ab initio molecular dynamics for weakly-interacting systems is developed 
%%for weakly-interacting systems. 
%%High efficiency of the methodology is achieved without sacrificing its accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable and correct dynamics even with imperfect forces. 
%%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with diffuse atom-centered basis sets. 
%Application of the method to liquid water demonstrates that the new AIMD method enables simulations of molecular systems on previously inaccessible time and length scales.

%The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation. 
%The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
%In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.

@@

%RZK important addition for a more detialed publication: review of a large family of fragmentation methods into the review following these guidelines.
%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix. This approach provides the most rigorous description of the electronic structure.

@@

ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems. %with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.

%To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to compensate errors in the forces~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}. 
%We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin equation. 
%We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.

The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals. First, while ALMOs are variationally optimized in both SCF steps, the occupied subspace defined in the first step must remain fixed during the second step to ensure convergence. Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$. In addition to these issues characteristic to ALMOs, the variational optimization in any AIMD method is never truly complete and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$. While these nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations, they accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure. 
Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$. 
%This however would dramatically increase the computational cost of the ALMO AIMD. %as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.

In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the complexity and cost of simulations low. 
%
In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that could be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced to zero \emph{systematically} by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$. %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
%RZK For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}. 
Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that are first written in terms of the unknown reference forces:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
% 
The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and will generate the correct Maxwell-Boltzmann distribution of velocities for a requested temperature $T$~\cite{a:Kubo-1986}. 
In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered. %and the microcanonical ensemble is sampled. 
Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to faster equilibration but more distorted results for dynamical properties.

@@

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD. %and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.

@@

%RZK: There is a problem with the integral equation below. The r.h.s. depends on $i$ but in the l.h.s. does not. If I remember correctly we evaluate the integral for molecules, not atoms. In liquid water all our molecules are the same and we do not have to deal with the issue of different Delta for different molecules. Eliminating the problem
In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle \delta f_{i\alpha} (0)  \delta f_{i\alpha} (t) \rangle dt
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$ .
%%
%\begin{align}
%\label{eq:eqipartition}
%\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
%\end{align}

@@

ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
% RZK: we do not need to specify details on gamma optimization since they are well-known.
Ricci-Ciccotti algorithm~\cite{RZK:10.1080/0026897031000108113} was used to integrate the Langevin equation. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
%to describe the interactions between the valence electrons and the ionic cores.

The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR). This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the reference radial distribution function (RDF) perfectly in Monte-Carlo simulations~\cite{a:almo-ls}. 
%
To measure the ability of the $R^{\Delta}$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
\caption{\label{fig:dynproperties} 
%RZK: "G_OO -> g_OO", "Intensity (au)" -> "Intensity", remove "DFT" in two legend labels 
Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) RDF, 
(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) IR spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

%The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
%at $\SI{302}{\K}$ and  --- RZK: It is customary to report TARGET temperature, not the acutal average temperature. 
%with the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
%All other system parameters were set identical to the ALMO system.

%RZK: 5K deviation from the target temperature is considered insignificant. Even with perfect forces, the actual temperature deviates from the target temperature by 5 K. This shows that our simulations are still too short to get the correct average. This is fine though.
Even with $\epsilon_{\text{SCF}} = 10^{-2}$, we stabilized the simulation at the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b). $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$. 
%RZK: factor of 3 bothers me. Check that this is not because we sum over 3 atoms in a molecule (or over 3 dimensions).
We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
%In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
%
Analysis of $\delta \vec{f_{i}}(t)$ shows that the error indeed resembles Gaussian white noise. The mean of the error is zero (black line in Figure~\ref{fig:randomforce}a). Its ACF decays rapidly (Figure~\ref{fig:randomforce}c) so that the errors can be considered uncorrelated on time scale of $\SI{50}{\fs}$. Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water. The main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the rapidly oscillating component of the error (Figure~\ref{fig:randomforce}b). We also confirmed that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ 
%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
(Figure~S1 in the Supplemental Material).

@@

%RZK addition for a longer article: extrapolation of WFs on both SCF steps is a crucial component of our fast algorithm.

To demonstrate the computational advantage of ALMO AIMD, we compared the average wall-time per MD step over 10 time steps for a variety of methods in Figure \ref{fig:strongscaling_log}.
In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically scaling OT method~\cite{a:2ndcpmd}.
As expected ALMO methods show clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems. 
%OT SCF scales approximately cubically, as expected.
For liquid water, strict localization of ALMOs lowers the crossover point between ALMO AIMD and cubically scaling methods to 256 molecules -- length scales routinely accessible with AIMD today. It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with a large diffuse basis set -- a particularly challenging case for all LS methods.
